---
title: "Correlations LOOR, DOOR and PORSCH outcome"
author: "Henri van Werkhoven"
date: "2026-07-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
spearman_fmt <- function(x){
  sprintf("%.3f (%.3f to %.3f)", x[1], x[2], x[3])
}
```

## Figure 1: scatter plot of mean LOOR outcome vs. DOOR outcome or components {.tabset}

Spearman correlations provided below each plot.

### A. DOOR

```{r}
plot_mean_loor_vs_door
```

Spearman correlation (95% CI): `r spearman_fmt(spearman_mean_loor_vs_door)`.

### B. PORSCH

```{r}
plot_mean_loor_vs_porsch
```

Spearman correlation (95% CI): `r spearman_fmt(spearman_mean_loor_vs_porsch)`.

### C. Organ failure

```{r}
plot_mean_loor_vs_OFD
```

Spearman correlation (95% CI): `r spearman_fmt(spearman_mean_loor_vs_OFD)`.

### D. Death

```{r}
plot_mean_loor_vs_death
```

Spearman correlation (95% CI): `r spearman_fmt(spearman_mean_loor_vs_death)`.

## Figure 2: Transition probabilities by DOOR rank {.tabset}

```{r, results='asis'}
for(j in 1:length(state_occupancy_by_door$x)){
  cat("\n\n### ", LETTERS[j], ": Rank ", state_occupancy_by_door$door[j], "\n\n")
  print(state_occupancy_by_door$x[[j]])
}
```

## Figure 3: Transition probabilities by PORSCH endpoint {.tabset}

```{r, results='asis'}
for(j in 1:length(state_occupancy_by_porsch$x)){
  cat("\n\n### ", LETTERS[j], ": ", as.character(state_occupancy_by_porsch$porsch[j]), "\n\n")
  print(state_occupancy_by_porsch$x[[j]])
}
```

## Figure 4: Transition probabilities by organ failure or death {.tabset}

```{r, results='asis'}
for(j in 1:length(state_occupancy_by_organfailure$x)){
  cat("\n\n### ", LETTERS[j], ": ", as.character(state_occupancy_by_organfailure$OFD[j]), "\n\n")
  print(state_occupancy_by_organfailure$x[[j]])
}
```

